A Kochen-Specker system has at least 21 vectors\\ {\small (extended abstract)}

### 2014-09-08 paper/main.tex
paper: update figures.  add attribution to Judith

Suppose there is a KS system and a non-contextual deterministic theory satisfying
the SPIN Axiom.
Then we color a point of this system~$0$,
whenever this theory predicts that the particle will deviate
if the spin is measured in the direction corresponding to that
point, and~$1$ otherwise.
Given two orthogonal points of the system,
we can find a third point orthogonal to both of them.
The SPIN axiom implies exactly one of them is colored~$1$, so they
cannot both be colored~$1$.
Similarly, given three pairwise orthogonal vectors in the system,
the SPIN axiom implies exactly one of them is colored~$1$.
Hence there would be a 010-coloring of the KS system, quod non.
Therefore a deterministic non-contextual theory cannot satisfy the
SPIN Axiom.

@@

\section{Embeddability}
\label{sec:emb}
\begin{wrapfigure}{r}{0.40\textwidth}
\begin{center}
\includegraphics[width=50mm]{../graphs/unemb-10-2.jpg}
\end{center}
\caption{One of the two minimal
        non-embeddable graphs
\label{fig:unemb-10-2}}
\end{wrapfigure}

Our computation has yielded a few hundred non-010-colorable graphs.
If we show one of them is embeddable, we have found a new KS system.
If we demonstrate all of them are not embeddable, we have
proven a lower bound on the size of a minimal KS system.

In~\cite{aow11}, Arends, Wampler and Ouaknine discuss several
computer-aided methods
to test embeddability of a graph.  None of these methods could decide
for all graphs considered, whether they were embeddable or not.
% TODO give a small description of their deficiencies
% TODO give intermediate computation.

@@

With this method, we have decided in an hour the embeddability
of every squarefree graph with minimal vertex order three of 13
vertices of less.
In particular:
\begin{comp}\label{comp:unemb20}
    Every squarefree graph of minimal vertex order three
    that is not 010-colorable
    of order less than or equal to 20
    contains, as a subgraph, one of the following three graphs:
    \begin{center}
        \includegraphics[width=120mm]{../graphs/unemb-base-20.jpg}
    \end{center}
    These three graphs are unembeddable.  The left and middle graph
    are the only minimal unembeddable squarefree graph.
\end{comp}
For the first graph, we have proven directly that it is unembeddable.
See Proposition~\ref{prop:unemb-10-2}.
For the second graph, we also have a similar direct proof. The third graph is shown to not be embeddable using our algorithm.

@@

We are grateful to prof.~McKay for discussing
the feasibility of certain graph restrictions.

>>>>

\begin{wrapfigure}{r}{0.30\textwidth}
\begin{center}
\includegraphics[width=40mm]{../graphs/ks31.pdf}
\end{center}
\caption{Conway's 31 vector Kochen-Specker system
\label{fig:conway31}}
\end{wrapfigure}
Suppose there is a KS system and a non-contextual deterministic theory satisfying
the SPIN Axiom.
Then we color a point of this system~$0$,
whenever this theory predicts that the particle will deviate
if the spin is measured in the direction corresponding to that
point, and~$1$ otherwise.
Given two orthogonal points of the system,
we can find a third point orthogonal to both of them.
The SPIN axiom implies exactly one of them is colored~$1$, so they
cannot both be colored~$1$.
Similarly, given three pairwise orthogonal vectors in the system,
the SPIN axiom implies exactly one of them is colored~$1$.
Hence there would be a 010-coloring of the KS system, quod non.
Therefore a deterministic non-contextual theory cannot satisfy the
SPIN Axiom.

@@

\section{Embeddability}
Our computation has yielded a few hundred non-010-colorable graphs.
If we show one of them is embeddable, we have found a new KS system.
If we demonstrate all of them are not embeddable, we have
proven a lower bound on the size of a minimal KS system.

In~\cite{aow11}, Arends, Wampler and Ouaknine discuss several
computer-aided methods
to test embeddability of a graph.  None of these methods could decide
for all graphs considered, whether they were embeddable or not.
% TODO give a small description of their deficiencies
% TODO give intermediate computation.
\label{sec:emb}
\begin{wrapfigure}{r}{0.30\textwidth}
\begin{center}
\includegraphics[width=50mm]{../graphs/unemb10-1-a.pdf}
\end{center}
\caption{One of the two minimal
        non-embeddable graphs
\label{fig:unemb-10-2}}
\end{wrapfigure}

@@

With this method, we have decided in an hour the embeddability
of every squarefree graph with minimal vertex order three of 13
vertices of less.
In particular:
\begin{comp}\label{comp:unemb20}
    Every squarefree graph of minimal vertex order three
    that is not 010-colorable
    of order less than or equal to 20
    contains, as a subgraph, one of the following three graphs:
    \begin{center}
        \includegraphics[width=30mm]{../graphs/unemb10-1.pdf} \qquad
        \includegraphics[width=30mm]{../graphs/unemb10-2.pdf} \qquad
        \includegraphics[width=30mm]{../graphs/unemb10-3.pdf}
    \end{center}
    These three graphs are unembeddable.  The left and middle graph
    are the only minimal unembeddable squarefree graph.
\end{comp}
For the first graph, we have proven directly that it is unembeddable.
See Proposition~\ref{prop:unemb-10-2}.
For the second graph, we also have a similar direct proof. The third graph is shown to not be embeddable using our algorithm.

@@

We are grateful to prof.~McKay for discussing
the feasibility of certain graph restrictions and
to Judith van Stegeren\footnote{\url{http://jd7h.com/}}
for drawing the figures in this paper.
